Projecting Global Filesystems into Local Containers for High-throughput Computing on HPC Resources

### 2019-01-24 paper.tex
Submitted Paper Version

\author{Tim Shaffer}
\email{tshaffe1@nd.edu}
\orcid{}
\affiliation{%
  \institution{University of Notre Dame}
  \city{Notre Dame}
  \state{Indiana}
  \postcode{46556}
}

\author{Nicholas Hazekamp}
\email{nhazekam@nd.edu}
\orcid{}
\affiliation{%
  \institution{University of Notre Dame}
  \city{Notre Dame}
  \state{Indiana}
  \postcode{46556}
}

\author{Jakob Blomer}
\email{jblomer@cern.ch}
\orcid{}
\affiliation{%
  \institution{CERN}
  \city{Geneva}
  \country{Switzerland}
}

\author{Douglas Thain}
\email{dthain@nd.edu}
\orcid{}
\affiliation{%
  \institution{University of Notre Dame}
  \city{Notre Dame}
  \state{Indiana}
  \postcode{46556}
}

@@

\begin{abstract}
Scientific applications often depend on large, complex software repositories containing libraries, tools, and frameworks in multiple versions and platforms to support many researchers.
Such repositories are distributed to hundreds of major computing
facilities around the world using a global filesystem like CVMFS.
However, not all HPC facilities permit execution nodes to
mount and access global network resources, instead encouraging
applications to make use of local filesystems and container
technologies.  In this work, we present a technique for efficiently.
\emph{projecting} a global filesystem into local resources.
This consists of creating minimal local containers that meet
the needs of specific jobs, balancing between the extremes
of having many small single-purpose containers and having one
large all-purpose containers.  We evaluate these techniques
using data from production instances of the CVMFS global filesystem
used by a wide variety of scientific collaborations.
\end{abstract}

@@

Global scientific experiment collaborations like ATLAS, CMS, LIGO,
and LSST make use of hundreds of thousands of cores spread across
many computing facilities around the world in order to achieve the massive scale needed simulate and analyze data from high bandwidth
scientific instruments.  In order to ensure consistency of software
running at many different sites, collaborations establish global
software repositories that publish the tested and trusted versions
of software to be used for production runs.  These repositories
can be quite large, consisting of thousands of software packages
consuming millions of files and terabytes of data.  Software
such as CVMFS~\cite{globalfs-cise-2015} is used to distribute
these complex software stacks around the world.

@@

First, for a given job, we must have a means of identifying the
subset of the global filesystem that is relevant for the current
execution.  This can be done either by dynamic tracing of the resources
actually used, or by static analysis of its dependency tree.
We describe Shrinkwrap, a tool that performs this analysis on
a Unix application, and produces a minimal container
that can be used to execute the application, or one closely related.
We analyze the efficiency of this tool on a selection of high
energy physics applications, and demonstrate storage efficiency of this technique, and measure the performance and storage efficiency.

@@

Scientific research has a large and rapidly growing need for computation.
Researchers have a number of options at their disposal to meet those needs,
including campus-local or organization-local resources,
cloud providers,
grid computing,
and large HPC sites.
Moving applications between these classes of resources can present problems,
as the assumptions made during development may no longer hold.
For example, organization-local resources and some grid sites commonly provide access to shared filesystems at each compute node.
In addition, there is some flexibility in system configuration to support an organization's primary research activity.
A new computing environment may be much less open to changes for a specific application,
especially large, multi-tenant resources such as HPC sites.
In these cases, it is left to scientists to adapt their application software to the available resources.

Many of our research collaborators have experienced significant frustration in preparing an existing scientific application to run on a new type of resource.
As a concrete example for this work,
we consider the challenges in running simulation and analysis workloads at CERN on HPC sites.
These applications assume several pieces of infrastructure to be in place and require some configuration and customization to the compute environment.
This is not an issue when running on Worldwide LHC Computing Grid~(WLCG),
which is a dedicated resource built to meet CERN's substantial computing needs.
The WLCG currently satisfies most of the computational needs of the LHC experiments at CERN.
With the upcoming high luminosity upgrades to the LHC,
the computational resources needed will will increase by at least an order of magnitude by 2026,
% On CERN's page it estimate 50-100X need, argonne had a report that was 10-20X
prompting researchers to turn to available high performance computing (HPC) resources to meet demand.
%In contrast with the WLCG's single slot submissions,
%HPC applications typically consist of parallel jobs using several cores up to a full machine for each submission.

Most HPC systems, however, have site-specific configuration and security requirements.
These can take the form of limited network availability at computation nodes,
limited container support,
and restricted access to privileged system features (such as FUSE support).
These new restrictions present challenges when porting existing scientific applications,
which expect required software to be readily available,
make a number of assumptions about how the system should be configured,
and may not be designed with tasks such as converting between image and container formats in mind.

In the case of the simulation and analysis workloads at CERN,
unrestricted network access and FUSE support are critical for providing global filesystem access.
Nearly all research software used at CERN assumes the availability of the CernVM File System (CVMFS),
a global filesystem built for distributing software and configuration data.
CMVFS provides a read-only, distributed file system using HTTP for network transport and extensive local
caching to rapidly distribute updates in software across the WLCG.
Dependence on CVMFS to distribute experiment software currently makes it impossible to take advantage of many HPC sites.
CVMFS requires unrestricted network access and is implemented as a FUSE module,
both of which sites often disallow for security and management reasons.

Some sites, such as NERSC,
have attempted workarounds for providing CVMFS,
such as creating large images several terabytes in size that contain all of the software stored in CVMFS or
maintaining local mirrors of CVMFS available via site-local shared filesystems.
However, both of these existing solutions require 
hours to days of attention from system administrators to deploy or update,
which leads to lag between the available software and the current version.
Additionally for systems that utilize large container images, large amounts of storage are needed to maintain different versions.

Since each LHC experiment job uses 
only a relatively small portion of CVMFS,
a more fine-grained approach to preparing container images would reduce the time spent creating images and allow for more precise control of the software versions in use.
This could also reduce the storage required at each worker node and the network bandwidth spent transferring images.
It is often possible to determine dependencies for particular applications statically by examining submission commands, setup scripts, etc.
This approach, while reliable, tends to be labor-intensive and tied to specific experiments and software stacks.
We therefore modified the reference implementation of CVMFS to allow us to capture the filesystem interactions of live HEP applications.
Using live tracing and static analysis,
we are able to determine exactly which parts of the \texttt{/cvmfs} file system tree CVMFS must be included in a container image for a particular job.
We developed a tool, Shrinkwrap, that can efficiently \emph{project} CVMFS (or a subset thereof) at a particular version into a traditional filesystem.
Such a \emph{projection} of CVMFS is suitable for inclusion in a static container image that does not depend on FUSE or network access.

While working with projected container images and Shrinkwrap,
we found that the cost of creating, distributing, and storing customized images for a large number of jobs and users creates unexpected challenges as well.
Shrinkwrap is designed to efficiently manage projections by deduplicating data in the filesystem.
Disk images, however, must contain complete copies of all included data.
The na\"{i}ve approach of creating a projection of CVMFS for each application results in frequent duplication of data and significant overhead in terms of computation and bandwidth.
Depending on the granularity of the projection,
image creation time can easily exceed individual job's actual compute time.
The alternative extreme, i.e.\ building a single large image that satisfies the requirements of all applications,
forces all jobs to use the same repository revision,
which can result in incompatibilities across applications.
Updates to the repository or changes in applications and requirements can again result in large amounts of overhead due to frequently creating, transferring, and storing these very large images.

Observing that jobs frequently have significant overlap in dependencies,
we designed a system to more efficiently manage container images on job submission.
First, we defined a metric for determining the similarity between projections.
This gives the possibility of identifying existing container images that could be re-used to fulfill a new request.
We also determined that in some cases,
it is less expensive to augment an existing, similar projection to allow a single container image to support multiple distinct applications.
We bring these tools together as a proof-of-concept submission system for making online decisions about whether to create, reuse, or augment container images to efficiently support high-throughput HEP jobs on HPC resources.

Compared to existing solutions,
Shrinkwrap created more compact projections than traditional methods such as rsync copies,
and allowed for more efficient updates to existing projections.
Based on the existing projects to produce static CVMFS container images and traces of benchmark applications,
we examined the sizes and structures of several projected container images.
Based on these properties,
we simulated the behavior of our image management system over a large number of job submissions.
We found that our approach can effectively manage the container images for a continuous stream of submissions with varying software dependencies and versions with less storage overhead than creating a new image per application.
In addition, our results show that applications can still take advantage of precise dependency and version specifications to reduce image creation and transfer overhead compared to working with large/complete mirrored copies of CVMFS.

\begin{figure}
\includegraphics[width=\linewidth]{drawings/methods.pdf}
\label{choices}
\caption{Global Filesystem Access Methods}
\end{figure}

@@

Modern scientific applications depend upon complex software
repositories that consist of terabytes of data and millions of files.
Unlike "traditional" high performance applications that are
single, statically-linked executables, modern applications are
agglomerations of software that consist of high level scripts,
multiple executables written in different languages, support
libraries that may be dynamically linked, and configuration and
data files loaded at runtime.

@@

Figure~\ref{choices} outlines several broad methods for
accessing such global filesystems.  First, the most desirable
method is to {\bf mount} the filesystem directly from worker
nodes, so that the necessary files can be accessed on demand.
While this is possible at sites that are especially configured
to support the LHC, it is often not possible at other HPC sites
that restrict general network access from worker nodes.
There are many social and technical reasons for these restrictions, but they generally boil down to either security (avoiding intrusions into worker nodes) or efficiency (discouraging high value workers from waiting on network access.)  As a result, many HPC sites are unable
to use global software repositories.

If that cannot be accomplished, then the next
alternative is to {\bf copy} the entire repository to
a shared parallel or distributed filesystem at the site.
This has the advantage that the filesystem is already
mounted and available at worker nodes.  The entire
repository must be duplicated before it can be used,
regardless of whether an individual job needs it all.
A more subtle problem is that most parallel filesystems are
designed for delivering very large files, and are not at all 
efficient at delivering deep directory trees that
consist of small files and large amounts of metadata.
When a large number of workers begin to use the software
simultaneously, the system can collapse from the metadata
load.  ~\cite{metafs-pdsw-2017}  (XXX CITE MORE HERE)

A {\bf container} is often used to reduce the metadata load
on a parallel filesystem.  If a complex software tree is
deployed into a container, the underlying filesystem will
treat the container as a single large file, and be more
efficient at delivering it to many worker nodes simultaneously.
However, if the entire repository (measured in TB) is placed into
the container, it is likely to exceed a number of practical
limits on container size.  Individual worker nodes may have limited
local disk space, or may be entirely diskless.  Even if the
large container fits, it is likely that a given job does not need
\emph{all} of the repository simultaneously, and so it is wasteful
to transfer unneeded data.

@@

\section{Defining Application Dependencies}

While the entirety of CVMFS is fairly large,
a single application would never use every piece of data.
Rather than attempting to provide a complete copy of CVMFS locally,
which entails significant storage costs,
we could instead provide a limited subset of the complete filesystem that is sufficient for a particular job.
This approach is more in line with traditional software packaging practices,
where software is provided as needed in the form of RPM files or another similar format.
Working with software packages is simpler in that dependencies are explicitly specified in the package file or database.
With CVMFS, no such definitive specification exists.
This issue is common to shared/global filesystems,
since users expect the complete contents to be available on demand and so devote little time to explicitly defining dependencies and other requirements.

In the case of CVMFS,
building an accurate 
specification of the files, libraries, and packages
needed for each experiment is the first step toward creating usable container images.
Such file specifications can be used when 
projecting CVMFS into realized filesystems.
The correctness and storage overhead of the projected filesystem hinges on the clarity and accuracy of the file specification.
File specifications can be derived from several
different sources, such as 
usage traces, static analysis, or package listings.
The end goal of any of these methods is to 
create a comprehensive list of needed files and structure.
The nature of global filesystems flexible behavior
further extends file specifications by requiring
not just the files and structure, but also what
version/revision was used to verify the contents as well.

\subsection{Creation of Specification}
File specifications can be created several ways,
with the obvious option being handwritten.
However, this quickly becomes unfeasible with
even a moderate number of requirements.
This leaves the need for an automate method to create
specifications.

\subsubsection{Tracing}
The first method is to create a sample trace for your
jobs. 
Creating a trace from a sample job allows 
for a listing of actual files used, which
can be more accurate than handwritten listings.
As these jobs are typically on the order of thousands
to millions of jobs, sampling a small subset
can provide good software coverage with little
additional overhead. % the computation is still used
The burden of creating these traces is
dependant on the method used,
which could be an LD\_PRELOAD library,
using Parrot to track used files, % I think parrot was mentioned previously
or building the tracing in to the application.

Tracing is attractive as it provides a consistent 
behavior between experiments and use cases.
In addition, it is easier to automate and sample arbitrary applications without assistance from each user.
This allows the solution to port between applications
quickly without having to delve in to the applications
specific configuration.
As previously mentioned, there are several methods for
obtaining a trace from a file.

Despite assumption that both CVMFS and FUSE will be limited
on the HPC site, this methods was used as it is reasonable to
allow for sampling of these jobs on WLCG. 
WLCG provides the flexibility of using CVMFS and FUSE, without
having to implement or provision a separate tracing infrastructure
at each site.
Additionally, as many of these jobs would have originally been 
destine to run on WLCG, the sampling sizes are unlikely to 
create significant overhead.
Using CVMFS for the tracing allows CVMFS's just-in-time deployment
of files to patch the environment and give an accurate view without
the risk of missing files.

\begin{figure}
\includegraphics[width=\linewidth]{curated/cdf/cdf.pdf}
\caption{The distribution of file size for each repository.
From this we can see that the large majority of files
accumulate in the kilobyte to megabyte range.}
\label{fig:cdf}
\if 0
lots of small files
table shows duplication
long tail of large files
different scales, similar usage across repos

\fi
\end{figure}

\subsubsection{Packages}
The second automated method is to use static
analysis for discover requirements.
This would allow the application to be analyzed
prior to running to determine what packages are
linked.
Static analysis can be done at several levels,
each providing different levels of control 
and flexibility.
This could be more language specific in the 
form of following import lines in python or
include lines in C.
The alternative is to use repository specific 
methods for setting up the environment,
such as bash scripts for pointing to correct 
software versions, or 
analyzing SCRAM\cite{DBLP:journals/corr/cs-OH-0306014} lines
to determine the configuration.
In the SFT CVMFS repository, where many core research packages are located,
dependencies are mapped for many packages using buildinfo files.
Chaining these buildinfo files together allow for a rough dependency
tree to be create, which was used to sample for \Cref{fig:picks}.
With either of these methods the issues
that arise come from bleeding edge or
experiment specific code that is not directly
handled by the configuration.
There is a significant mix between languages,
configurations methods, and software organization
practices even within a single repository,
which compounds the above issues in this method.

An combination of tracing and static analysis is the
use of package listings. 
Package listings provide a high level indication
of the software needed, and allow for 
flexibility on the specifying user about the
exact files needed.
Using high level packages provides a method
for handwritten/tuned specifications that 
include all aspects of the package.
The downside of using packages is that in
their flexible, forgiving specification
they can be bloated and include unnecessary
files and data.
Packages also require that there is consistent
software organization, which often differs
between repositories.

\begin{figure}
\includegraphics[width=\linewidth]{curated/closure/closure.pdf}
\caption{Difference in number of packages and size when accounting for dependency based closure in selection.}
\label{fig:picks}
\if 0
you get more than you ask for
smaller requirements have relatively large responses
non-linear, so bigger ones have more duplicates
appraoches repo size in limit
merging is subject to drop-off(more duplication)
\fi
\end{figure}

\subsection{Unique Consistency Requirements}
CVMFS differs from traditional POSIX filesystems in that previous versions of all content remain available.
A researcher may specify the precise revision under which an application is known to run.
Thus the same path may refer to two completely different files for a pair of applications.
This is beneficial from the standpoint of reliability and reproducibility,
but complicates the process of projection.
A file specification needs to include enough information
to establish the directory structure and verify the contents at a specific revision.
This can be done in several ways, 
such as computing the checksum of contents,
using global revision numbers, 
or running verification software.
Computed checksums provide a system agnostic approach
to verification by allowing the checksum to be computed
at specification and verified before use.
Computed checksums provide a
consistency guarantee to the byte level,
but are often expensive at runtime for large
production systems.
This approach is alleviated in systems such as 
Git and CVMFS as the contents' checksums are 
computed at ingestion into the repository.
This method relies on the underlying system computing
and providing this information for verification.

\if 0
\begin{table*}[t]
\begin{center}
\begin{tabular}{|c|r r r|r r r|}
\hline
CVMFS & & ls -lr & & & Shrinkwrap & \\
Repository & Total Size (MB) & inode Count & Time to walk & 
  Data Files & Deduplicated Data Files & Deduplicated Data (MB) \\ \hline
%alice-ocdb & 482,642,307,428 & 1,621,661 & 2:22 &
% 1,621,282 & 1,595 & 24,433,131,203\\
%alice & 3,980,875,747,986 & 27,322,350 & 51:53 & 
%  797,404 & 23,996,541 & 2,098,509,015,172\\
alice & 3,980,875 & 27,322,000 & 51:53 & 
  797,000 & 23,997,000 & 2,098,509\\
%atlas-condb & 719,548,941,240 & 11,781 & 0:04 &
%  10,940 & 229 & 471,492,086 \\
%atlas & 6,195,013,689,056 & 120,598,601 & 3:48:07 &
% 8,823,179 & 85,554,444 & 3,760,712,059,548 \\
atlas & 6,195,013 & 120,599,000 & 3:48:07 &
 8,823,000 & 85,554,000 & 3,760,712 \\
%boss & 120,048,527,315 & 2,418,025 & 4:18 &
% 256,061 & 1,868,192 & 72,301,955,666\\
boss & 120,048 & 2,418,000 & 4:18 &
 256,000 & 1,868,000 & 72,302\\
%cms & 581,789,406,847 & 26,224,567 & 1:07:30 &
% 18,667,178 & 101,448,867 & *2,242,658,825,701\\
cms & 581,789 & 26,225,000 & 1:07:30 &
 18,667,000 & 101,449,000 & *2,242,658\\
%geant4 & 297,571,283,648 & 7,128,024 & 10:36 &
% 323,280 & 6,219,891 & 68,381,969,660 \\
geant4 & 297,571 & 7,128,000 & 10:36 &
 323,000 & 6,220,000 & 68,382 \\
%grid & 30,642,826,259 & 714,235 & 1:17 &
% 141,622 & 583,923 & 18,804,003,040 \\
grid & 30,643 & 714,000 & 1:17 &
 142,000 & 584,000 & 18,804 \\
%lhcb & 2,134,893,079,440 & 33,736,802 & 1:01:26 &
% 10,032,339 & 20,406,235 & 368,070,277,634 \\
lhcb & 2,134,893 & 33,737,000 & 1:01:26 &
 10,032,000 & 20,406,000 & 368,070 \\
%na61 & 5,222,967,734 & 25,602 & 0:02 &
% 8,609 & 15,001 & 2,425,507,431 \\
na61 & 5,222 & 26,000 & 0:02 &
 9,000 & 15,000 & 2,426 \\
%sft & 3,996,527,275,237 & 128,608,792 & 4:17:58 &
% 6,821,699 & 63,210,139 & 2,016,609,810,573 \\ 
sft & 3,996,527 & 128,609,000 & 4:17:58 &
 6,822,000 & 63,210,000 & 2,016,610 \\ 
 \hline
\end{tabular}
\caption{CVMFS repository sizes.}
\label{tab:repo-sizes}
This table shows the size, number of inodes (rounded to nearest thousand), and the amount
of duplicate data that is eliminated using CVMFS.
\end{center}
\end{table*}

\begin{table}
\begin{center}
\begin{tabular}{|c|r r r|}
\hline
Repository & Total Size & inode Count & Deduplicated \\
 & (GB) & (K) & Data (GB) \\ \hline
alice & 3,981 & 27,322 & 2,099 \\
atlas & 6,195 & 120,599 & 3,761 \\
boss & 120 & 2,418 & 72\\
cms & 582 & 26,225 & *2,243\\
geant4 & 298 & 7,128 & 68 \\
grid & 31 & 714 & 19 \\
lhcb & 2,135 & 33,737 & 368,070 \\
na61 & 5 & 26 & 2 \\
sft & 3,997 & 128,609 & 2,017 \\ 
 \hline
\end{tabular}
\caption{CVMFS repository sizes.}
\label{tab:repo-sizes}
This table shows the size, number of inodes (rounded to nearest thousand), and the amount
of duplicate data that is eliminated using CVMFS.
\end{center}
\end{table}

\section{Realization of CVMFS Projection}

\subsection{Definition of Projection}

>>>>

\author{Tim Shaffer$^\star$, Nicholas Hazekamp$^\star$, Jakob Blomer$^\dagger$, and Douglas Thain$^\star$}
\email{tshaffe1@nd.edu, nhazekam@nd.edu, jblomer@cern.ch, dthain@nd.edu}
%\orcid{}
\affiliation{%
  \institution{$^\star$ University of Notre Dame   $^\dagger$ European Laboratory for Particle Physics (CERN)}
%  \city{Notre Dame}
%  \state{Indiana}
%  \postcode{46556}
}

%\author{Jakob Blomer}
%\email{jblomer@cern.ch}
%\orcid{}
%\affiliation{%
%  \institution{CERN}
%  \city{Geneva}
%  \country{Switzerland}
%}

\if 0  
\author{Nicholas Hazekamp}
\email{nhazekam@nd.edu}
\orcid{}
\affiliation{%
  \institution{University of Notre Dame}
  \city{Notre Dame}
  \state{Indiana}
  \postcode{46556}
}

\author{Douglas Thain}
\email{dthain@nd.edu}
\orcid{}
\affiliation{%
  \institution{University of Notre Dame}
  \city{Notre Dame}
  \state{Indiana}
  \postcode{46556}
}  
\fi

@@

\begin{abstract}
Scientific applications often depend on large, complex software repositories containing libraries, tools, and frameworks in multiple versions and platforms to support many researchers.
Such repositories are distributed to hundreds of major computing
facilities around the world using a global filesystem like CVMFS.
However, not all HPC facilities permit execution nodes to
mount and access global network resources, instead encouraging
applications to make use of local filesystems and container
technologies.  In this work, we present a technique for efficiently
\emph{projecting} a global filesystem into local resources.
This consists of creating minimal local containers that meet
the needs of specific jobs, balancing between the extremes
of having many small single-purpose containers and having one
large all-purpose containers.  We evaluate these techniques
using data from production instances of the CVMFS global filesystem
used by a wide variety of scientific collaborations.
\end{abstract}

@@

Global scientific experiment collaborations like ATLAS, CMS, LIGO,
and LSST make use of hundreds of thousands of cores spread across
many computing facilities around the world in order to achieve the 
massive scale needed to simulate and analyze data from high bandwidth
scientific instruments.  In order to ensure consistency of software
running at many different sites, collaborations establish global
software repositories that publish the tested and trusted versions
of software to be used for production runs.  These repositories
can be quite large, consisting of thousands of software packages
containing millions of files and terabytes of data.  Software
such as CVMFS~\cite{globalfs-cise-2015} is used to distribute
these complex software stacks around the world.

@@

First, for a given job, we must have a means of identifying the
subset of the global filesystem that is relevant for the current
execution.  This can be done either by dynamic tracing of the resources
actually used, or by static analysis of its dependency tree.
Using the subset or specification from either method, a projection
then captures the job's software stack.
From this captured software, a minimal container can be
constructed to allow for more efficient storage, transfer, 
and use. We describe Shrinkwrap, 
a tool that produces a minimal software stack 
that can be used to execute the application, or ones closely related.
We analyze the size and make up of global filesystem repositories,
specifically in CVMFS, to show why existing solutions fall short.

@@

\begin{figure}
\includegraphics[width=\linewidth]{drawings/methods.pdf}
\caption{Global Filesystem Access Methods}
\label{fig:choices}
\end{figure}

@@

Modern scientific applications depend upon complex software
repositories that consist of terabytes of data and millions of files.
Unlike ``traditional'' high performance applications that are
single, statically-linked executables, modern applications are
agglomerations of software that consist of high level scripts,
multiple executables written in different languages, support
libraries that may be dynamically linked, and configuration and
data files loaded at runtime.~\cite{casestudy-poster-chep-2015}

@@

\Cref{fig:choices} outlines several broad methods for
accessing such global filesystems.  The simplest
method is to {\bf mount} the filesystem directly from worker
nodes, so that the necessary files can be accessed on demand.
While this is possible at sites that are especially configured
to support the LHC, it is often not possible at other HPC sites
that restrict general network access from worker nodes.
There are many social and technical reasons for these restrictions, but they generally boil down to either security (avoiding intrusions into worker nodes), efficiency (discouraging high value workers from waiting on network access), or configuration (kernel modules are not available to perform the mount.)  As a result, many HPC sites are unable
to use global filesystems of this form.

If that cannot be accomplished, then the next
alternative is to {\bf copy} the entire repository to
a shared parallel or distributed filesystem at the site.
This has the advantage that the filesystem is already
mounted and available at worker nodes.  The entire
repository must be duplicated before it can be used,
regardless of whether an individual job needs it all.
A more subtle problem is that most parallel filesystems are
designed for delivering very large files, and are not at all 
efficient at delivering deep directory trees that
consist of small files and large amounts of metadata,
which describes these repositories as seen in \Cref{fig:cdf}.
When a large number of workers begin to use the software or perform similar metadata operations
simultaneously, the system can collapse from the metadata
load.~\cite{metafs-pdsw-2017,spindle,torrent,exa}

A {\bf container} is often used to reduce the metadata load
on a parallel filesystem.  If a complex software tree is
deployed into a container, the underlying filesystem will
treat the container as a single large file, and be more
efficient at delivering it to many worker nodes simultaneously.
However, if the entire repository (measured in TB) is placed into
the container, it is likely to exceed a number of practical
limits on container size.  Individual worker nodes may have limited
local disk space, or may be entirely diskless.  Even if the
large container fits, it is likely that a given job does not need
\emph{all} of the repository simultaneously, and so it is wasteful
to transfer unneeded data.
This concept is a driving influence on projects like Slacker\cite{194430}.

@@

\begin{figure}[h]
\includegraphics[width=\linewidth]{curated/cdf/cdf.pdf}
\caption{Distribution of file sizes for each repository.}
\label{fig:cdf}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%

\section{Creating Projections}

Having defined a projection of a global filesystem as a subset of the filesystem stored locally,
the first step toward creating projections of a global repository is deciding which components to include.
Scientific applications use a wide variety of files, libraries, and packages,
and tend not to respond well to missing components.
When directly mounting a global filesystem or making a local copy,
the entire contents of the filesystem are assumed to be available,
even if any single job only uses a small portion of the total data.
As such, there may not be a convenient and complete description of which pieces of data are required.
Attempting to use a projection based on an incomplete specification leads to potentially intermittent job failures and wasted computing resources.
On the other hand,
including too broad a selection of data in a projection is undesirable as it increases overhead due to extraneous data in the container images.

The simplest way to create a specifications for a projection is manually listing each piece to include.
However, this quickly becomes unfeasible with even a moderate number of software component requirements.
In addition, manual specifications do not necessarily reflect the actual requirements of an application.
Changes to software or simple mistakes can result in incomplete specifications,
and include unnecessary files and data that result in bloated container images.
Manual specifications also practically require consistent, uniform software organization.
This is not a safe assumption within a single repository,
to say nothing of supporting multiple software sources.
This leaves the need for an automated method to create
specifications, such as 
 static analysis or runtime tracing.
To create a comprehensive list of required files and data,
it may be necessary to use a combination of these approaches.

\subsection{Static Dependency Analysis}

Static dependency analysis takes advantage of metadata, language, and application-specific knowledge to
discover requirements.
This technique allows applications to be analyzed
prior to running to determine what dependencies from the global filesystem are required.
Static analysis can be informed by several sources of information,
each with varying levels of detail and flexibility.
Language-specific analysis of software components might inspect the \texttt{import} statements in Python or
\texttt{include} directives in C.
These approaches only apply to specific languages or frameworks,
and require some knowledge about how the application is structured.
The software stack used for experiments can also provide information about dependencies in the way it sets up the working environment.
This form of static analysis might involve examining shell scripts for selecting  correct 
software versions, or 
analyzing any number of systems built for 
declaring dependencies, such as 
modules\cite{McLay:2011:BPD:2063348.2063360},
Spack\cite{Gamblin:2015:SPM:2807591.2807623},
VC3-Builder\cite{tovar-ic2e-2018},
or Nix\cite{Dolstra2008NixOSAP},
to determine the configuration.
Additionally it is not uncommon for experiments 
to design and use their own methods such as
SCRAM\cite{DBLP:journals/corr/cs-OH-0306014},
which is used in LHC experiments.
This kind of approach is highly tailored to specific applications or software frameworks and is difficult to generalize to other uses.

In the SFT CVMFS repository, where many of the core research packages at CERN are hosted,
dependency information is included in build metadata associated with each package.
This metadata gives a convenient way to construct a dependency
tree of the entire repository.
We constructed such a dependency tree for the SFT repository to use in the simulations discussed later.
\Cref{fig:picks} gives an indication of the structure and sizes of software components in the repository.
To construct that figure,
we started by repeatedly making random package selections of varying sizes and measuring the total sizes of the selections.
The two lines near the bottom of the graph show these counts and sizes.
Then for each selection of packages,
we used the extracted package metadata to recursively include each package's dependencies.
This recursive resolution simulates the procedure that a program might perform when starting up and loading libraries.
After finding the complete set of dependencies required by each selection,
we plotted the resulting package counts and sizes.
We repeated this procedure 100 times for each choice of specification size,
up to specifications of 1,000 packages.

The plot in \Cref{fig:picks} shows the medians of the resulting measurements.
We observe of course that selection size increases uniformly along the $x$ axis.
The on-disk size of the selections appears to increase proportionally to selection size.
Recursively including the dependencies of these same selections,
however, results in a significant increase in number of packages and size on disk.
For small selections (less than 100~packages),
the complete images with dependencies might contain 5x as many packages as requested,
with storage increasing proportionally.
With larger selections, however,
this increase becomes less prominent.
This non-linear change is the result of the tree structure of the software dependencies.
There are a number of core components that are transitive dependencies of a large number of packages.
These common dependencies are only included once,
so subsequent selections that depend on them do not add them again.
As the number of selections increases,
this curve will approach the total repository size.
Also note that the non-linear increase in size with increasing selection size means that the size of combining two selections of 100 packages (and their dependencies) is not as simple as adding them together.
Due to duplication, we would expect the resulting combined size to be lower,
as if we were choosing 200 packages from the start.

\begin{figure}
\includegraphics[width=\linewidth]{curated/closure/closure.pdf}
\caption{Difference in number of packages and size when accounting for dependency based closure in selection.}
\label{fig:picks}
\if 0
you get more than you ask for
smaller requirements have relatively large responses
non-linear, so bigger ones have more duplicates
approaches repositories size in limit
merging is subject to drop-off(more duplication)
\fi
\end{figure}

\if 0
\begin{figure}
    \includegraphics[width=\linewidth]{curated/lxplus_dist/usage_distribution_cdf.pdf}
    \caption{Usage distribution pulled from CERN LXPlus interactive machines.}
    \label{fig:usage_distribution}
\end{figure}

spread across everyone, lot of commonality/overlap
few with very high usage, very large number of rarely used
lot not observed
\fi

\subsection{Runtime Tracing}
Rather than examining the static data and metadata associated with a software collection,
we can use runtime tracing to identify the dependencies used during a particular job.
This trace of a sample job gives an exact listing of the files used,
which is more accurate and reliable than handwritten listings.
A potential drawback is that tracing does not determine every possible dependency,
only those used during a particular sample.
This can be problematic for non-deterministic or data-dependent workloads.
As such, it is preferable to trace a number of jobs.
Since scientific applications routinely consist of thousands
to millions of jobs, sampling a small subset
can provide good software coverage with little
additional overhead. % the computation is still used
The performance cost of creating these traces is
dependant on the method used.
A tracer library based on \texttt{LD\_PRELOAD} can be extremely lightweight but require modification to the execution environment.
More invasive methods include using Parrot to track used files, % I think parrot was mentioned previously
or building the tracing as part of the application.

An important advantage of runtime tracing techniques is that they can function reliably across multiple languages and applications.
This makes it is easier to automate sampling of unknown applications without assistance from each user,
since there is no need to delve in to the particular application
design and configuration.

As part of this work, runtime tracing was introduced
into the CVMFS FUSE client and can be activated for sampling.
When tracing an application,
each filesystem access is recorded and added to the trace log.
Tracing through CVMFS is convenient as it does not require changes to applications,
and the FUSE client is widely deployed across compute sites.
Users have the option of collecting this trace information themselves by re-configuring the client.
For applications running on the Worldwide LHC Computing Grid~(WLCG),
it would also be possible to instrument a fraction of runs to build a profile of a given task.
This introduces the possibility of automatically collecting traces to use when porting applications to more restrictive environments.

\subsection{Building Projections}

\begin{table}
\begin{center}
\begin{tabular}{|c|r r r|}
\hline
Repository & Total Size & inode Count & Deduplicated \\
 & (GB) & (K) & Data (GB) \\ \hline
alice & 3,981 & 27,322 & 2,099 \\
atlas & 6,195 & 120,599 & 3,761 \\
%boss & 120 & 2,418 & 72\\
cms & 582 & 26,225 & *2,243\\
%geant4 & 298 & 7,128 & 68 \\
grid & 31 & 714 & 19 \\
lhcb & 2,135 & 33,737 & 368,070 \\
%na61 & 5 & 26 & 2 \\
sft & 3,997 & 128,609 & 2,017 \\ 
 \hline
\end{tabular}
\caption{CVMFS repository sizes.}
\label{tab:repo-sizes}
{\small \it This table shows the size, number of inodes (rounded to nearest thousand), and the amount
of duplicate data that is eliminated using CVMFS.}
\end{center}
\end{table}

Several existing utilities such as UnCVMFS and 
rsync have been used to create projections of CVMFS.
UnCVMFS is a utility that projects an entire
CVMFS repository, at a set revision, into the local
filesystem. 
UnCVMFS utilizes CVMFS's efficient, content-based storage
to deduplicate the underlying data, but requires
entire repositories to be projected.
When used with several repositories for large projects,
unCVMFS can easily consume terabytes of disk space for a single projection.
In addition, it is expensive to update to different revisions of a repository using unCVMFS,
which can create headaches when working with different experiments and analyses that may have conflicting revision requirements.

Rsync allows for file lists to specify and download
subsets of CMVFS repositories, but because it
was designed for general filesystem synchronization
rsync is not aware of the internal organization of CVMFS.
This leads to slow operations and unnecessary duplication of data,
especially when creating multiple projections.

We built the CVMFS Shrinkwrap utility to create projections of CVMFS with better efficiency and finer control than is possible either unCVMFS or rsync individually.
Shrinkwrap uses the CVMFS's client interface to
create and update CVMFS projections.
Shrinkwrap uses a CVMFS configuration to specify the
revision of the repository, a file-specification list
to determine the repository subset to project, and
deduplicates the data in a user specified location.
Shrinkwrap can be configured to share the data
repository between several repositories and even
projections of the same repositories.
This allows for multiple realized projections
to coexist, with the desired projection
mounted for use.
In \Cref{figure:shrinkwrap-arch} you can see an outline
of the design of the Shrinkwrap utility.

\begin{figure}[h]
\includegraphics[width=\columnwidth]{drawings/shrinkwrap-structure.png}
\caption{Operation of Shrinkwrap}
\label{figure:shrinkwrap-arch}
\end{figure}

\subsection{Building Container Images}
